Convolutional Neural Networks and Volcano Plots: Screening and Prediction of Two-Dimensional Single-Atom Catalysts for CO\textsubscript{2} Reduction Reactions

### 2024-01-14 paper/arxiv/main_4.3_validation.tex
name formal names for submission
format fixes
make subfig symbol bold

\begin{figure}[htbp]
    \centering
    \includegraphics[width=\textwidth]{main_fig4_validation.png}
    \caption{\textbf{Validation of CNN Predictions.}
    a-c, The occlusion experiment results for (a)Cr $d_{x^2-y^2}$ orbital of Cr@g-C\textsubscript{3}N\textsubscript{4}, (b)Co $d_{xy}$ orbital of Co@g-C\textsubscript{3}N\textsubscript{4} and (c)Cu $d_{xy}$ orbital of Cu@g-C\textsubscript{3}N\textsubscript{4}.
    Occlusion experiments were conducted on the initial states of CO adsorption process.
    d-f, Real-space wavefunction visualization for (d)Cr $d_{x^2-y^2}$ orbital of Cr@g-C\textsubscript{3}N\textsubscript{4}, (e)Co $d_{xy}$ orbital of Co@g-C\textsubscript{3}N\textsubscript{4} and (f)Cu $d_{xy}$ orbital of Cu@g-C\textsubscript{3}N\textsubscript{4}.
    Wavefunction visualizations were performed on the final states of CO adsorption process.
    The yellow and aqua colors represent areas of positive and negative Kohn-Sham orbital coefficients, respectively.
    g-j, Occlusion experiment results for (g)$d_{xy}$ orbital and (h)$d_{xz}$ orbital for Co of Co@g-C\textsubscript{3}N\textsubscript{4}.
    COHP analysis for (i)$d_{xy}$ orbital and (j)$d_{xz}$ orbital for Co of Co@g-C\textsubscript{3}N\textsubscript{4}, performed between Co and all adsorbate atoms in the final states of CO adsorption process.}
    \label{main_fig4:validation}
\end{figure}

Historically, interpretability has posed challenges for deep learning methods, often rendering their inner workings akin to black boxes \cite{zhang2018interpretable, zhang2018visual, savage2022breaking}.
The occlusion technique, widely utilized in Computer Vision \cite{zeiler2014visualizing, kortylewski2020combining, wang2020robust}, involves systematically masking sections of the input matrix to assess their impact on the final prediction score.
In our work, we introduce an orbitalwise occlusion method to discern individual orbital contribution to adsorption energy within the CNN model, drawing inspiration from Fung's pioneer integration of this technique into eDOS analysis \cite{fung2021machine}.
In this work, occlusion experiments involve orbitalwise masking specific eDOS, which are then fed into the CNN model to observe resulting disturbance in adsorption energy.
A visual representation of this process can be found in Supplementary \cref{si_fig21:occlusion}. Conceptually, this experiment simulates the effect of shielding electronic states, allowing investigation of the potential contribution to adsorption energy from these states.
Additionally, we attempted to extract the planewave coefficients of Kohn-Sham orbital for wavefunction visualization.
This validation step was undertaken to confirm the spatial distribution of the orbitals pinpointed by the occlusion experiments.

>>>>

% Figure 4: Validation of CNN Predictions
\begin{figure}[htbp]
    \centering
    \includegraphics[width=\textwidth]{main_fig4_validation.png}
    \caption{\textbf{Validation of CNN Predictions.}
    \textbf{a}-\textbf{c}, The occlusion experiment results for (\textbf{a})Cr
    $d_{x^2-y^2}$ orbital of Cr@g-C\textsubscript{3}N\textsubscript{4},
    (\textbf{b})Co $d_{xy}$ orbital of Co@g-C\textsubscript{3}N\textsubscript{4} and
    (\textbf{c})Cu $d_{xy}$ orbital of Cu@g-C\textsubscript{3}N\textsubscript{4}.
    Occlusion experiments were conducted on the initial states of CO adsorption process.
    \textbf{d}-\textbf{f}, Real-space wavefunction visualization for (\textbf{d})Cr
    $d_{x^2-y^2}$ orbital of Cr@g-C\textsubscript{3}N\textsubscript{4}, (\textbf{e})Co
    $d_{xy}$ orbital of Co@g-C\textsubscript{3}N\textsubscript{4} and (\textbf{f})Cu
    $d_{xy}$ orbital of Cu@g-C\textsubscript{3}N\textsubscript{4}.
    Wavefunction visualizations were performed on the final states of CO adsorption process.
    The yellow and aqua colors represent areas of positive and negative
    Kohn-Sham orbital coefficients, respectively.
    \textbf{g}-\textbf{j}, Occlusion experiment results for (\textbf{g})$d_{xy}$ orbital
    and (\textbf{h})$d_{xz}$ orbital for Co of Co@g-C\textsubscript{3}N\textsubscript{4}.
    COHP analysis for (\textbf{i})$d_{xy}$ orbital and (\textbf{j})$d_{xz}$ orbital for
     Co of Co@g-C\textsubscript{3}N\textsubscript{4}, performed between Co and
     all adsorbate atoms in the final states of CO adsorption process.}
    \label{main_fig4:validation}
\end{figure}

Historically, interpretability has posed challenges for deep learning methods, often rendering their inner workings akin to black boxes \cite{zhang2018interpretable, zhang2018visual, savage2022breaking}.
The occlusion technique, widely utilized in Computer Vision \cite{zeiler2014visualizing, kortylewski2020combining, wang2020robust}, involves systematically masking sections of the input matrix to assess their impact on the final prediction score.
In our work, we introduce an orbitalwise occlusion method to discern individual orbital contribution to adsorption energy within the CNN model, drawing inspiration from Fung's pioneer integration of this technique into eDOS analysis \cite{fung2021machine}.
In this work, occlusion experiments involve orbitalwise masking specific eDOS, which are then fed into the CNN model to observe resulting disturbance in adsorption energy.
A visual representation of this process can be found in Supplementary \cref{supp_fig21:occlusion}. Conceptually, this experiment simulates the effect of shielding electronic states, allowing investigation of the potential contribution to adsorption energy from these states.
Additionally, we attempted to extract the planewave coefficients of Kohn-Sham orbital for wavefunction visualization.
This validation step was undertaken to confirm the spatial distribution of the orbitals pinpointed by the occlusion experiments.

### 2024-01-14 paper/arxiv/main_4.4_shifting_exp.tex
format fixes
make subfig symbol bold

\begin{figure}[htbp]
    \centering
    \includegraphics[width=\textwidth]{main_fig5_shifting.png}
    \caption{\textbf{Shifting Experiments.}
    Impact of orbitalwise eDOS shifting on CO adsorption energy, as predicted by the CNN model,
    for single metal atom catalysts supported on g-C\textsubscript{3}N\textsubscript{4}.
    The disturbances caused by shifting of (a)entire $d$, (b)$d_{xy}$, (c)$d_{yz}$,
    (d)$d_{z^2}$, (e)$d_{xz}$ and (f)$d_{x^2-y^2}$ orbitals are illustrated.
    The shifting step size corresponds to the energy resolution of the eDOS, which is 0.005 eV in this study.
    A positive eDOS shift indicates a shift towards higher energy levels, and vice versa.}
    \label{main_fig5:shifting}
\end{figure}

>>>>

% Figure 5: Shifting Experiments
\begin{figure}[htbp]
    \centering
    \includegraphics[width=\textwidth]{main_fig5_shifting.png}
    \caption{\textbf{Shifting Experiments.}
    Impact of orbitalwise eDOS shifting on CO adsorption energy, as predicted by the CNN model,
    for single metal atom catalysts supported on g-C\textsubscript{3}N\textsubscript{4}.
    The disturbances caused by shifting of (\textbf{a})entire $d$, (\textbf{b})$d_{xy}$,
    (\textbf{c})$d_{yz}$, (\textbf{d})$d_{z^2}$, (\textbf{e})$d_{xz}$ and
    (\textbf{f})$d_{x^2-y^2}$ orbitals are illustrated.
    The shifting step size corresponds to the energy resolution of the eDOS,
    which is 0.005 eV in this study.
    A positive eDOS shift indicates a shift towards higher energy levels, and vice versa.}
    \label{main_fig5:shifting}
\end{figure}
